---
title: "simple document"
output: github_document
date: "2023-09-19"
---



```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Let's import the `FAS_litters.csv` csv using a relative path.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

litters_df = 
  janitor::clean_names(litters_df) ##to clean up the names of the columns---convert spaces and uppercases to all lower cases and underscores(typical r name writing)


```

Import the same data set using an absolute path.

```{r}
litters_df_abs= 
  read_csv("~/Desktop/data_wrangling_i/data/FAS_litters.csv")

litters_df_abs = 
  janitor::clean_names(litters_df_abs) 

```

Import the `FAS_pups.csv` file using a relative path.

```{r}
pubs_df = 
  read_csv("data/FAS_pups.csv")

pubs_df = 
  janitor::clean_names(pubs_df)
```

Import the same data set using an absolute path.

```{r}
pubs_df_abs= 
  read_csv("~/Desktop/data_wrangling_i/data/FAS_pups.csv")

pubs_df_abs = 
  janitor::clean_names(pubs_df_abs) 
```

## Look at data

```{r}
litters_df

head(litters_df)

tail(litters_df)
```

You can use `View`

```{r, eval=FALSE} 
view(litters_df)
##use eval = FALSE so it wouldn't include in the r markdown
```

look at data summary: 2 ways 

```{r}
str(litters_df)
skimr::skim(litters_df) #with mean sd hist missing numbers
```


## Options in `read_*`

```{r}
##litters_df = 
 ## read_csv("data/FAS_litters.csv",
           ##skip = 10, col_names = FALSE) ##skip is used to skip rows when first x rows are explaining the data variables 
```


look at NA values

```{r}
litters_df =
  read_csv("data/FAS_litters.csv",
           na = c("NA", 19))
litters_df
```


column types: you can control the type of variables by changing num to char or otherwise

```{r}
litters_df =
  read_csv("data/FAS_litters.csv",
           col_types = 
             cols(
              `GD0 weight` = col_character() 
             ))
```


## Other File Types

Import a xlsx file first: first load a readxl package to read the excel files.


```{r}
mlb_df = 
  read_excel("data/mlb11.xlsx")
```

Import a SAS file: load a haven package in library first.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")
```


## Base R...

don't do this. 

```{r, eval = FALSE}
litters_df =
  read.csv("data/FAS_litters.csv")
```




## Exporting data 

We have code that 

```{r}
litters_df_cleaned = 
  read_csv("data/FAS_litters.csv")

litters_df_cleaned = 
  janitor:: clean_names(litters_df_cleaned)

write_csv(litters_df_cleaned, "data/litters_cleaned.csv")
```

